---
title: "Interactive Figures: RNA-seq and prospects for obtaining non-model plant transcriptomes"
author: "Alexa Tyszka"
date: "`r Sys.Date()`"
output: html_document
#runtime: shiny

---

```{r setup, include=FALSE}
#rmarkdown::render(input = "botany-23-supp/botany-23-supp.Rmd")
knitr::opts_chunk$set(echo = T)
library(raster)
library(plotly)
library(ggplot2)
library(geodata)
library(viridis)
library(tidyterra)
library(terra)
library(sf)
w <- read_sf(dsn= "World_Countries_Generalized/World_Countries_Generalized.shp")
buttonstoremove <- c('zoom2d', 'pan2d', "toImage", "select2d", "lasso2d", "autoScale2d", "hoverClosestCartesian", "hoverCompareCartesian")
```


## Figure 1a: Total global vascular plant species richness



```{r fig1a, echo=FALSE, warning=FALSE, out.width='375px', out.height="667px"}

  rich_whole = raster("rich_whole_4.tif")
rwd <- as.data.frame(rich_whole, xy=T)
rwd$count <- rwd$scrubbed_species_binomial 


p  <- ggplot(rwd, plot.margin = margin(0, 0, 0, 0, "cm")) +
   geom_sf(data=w, fill=NA)+
    #geom_spatraster(data = rich, aes(fill = scrubbed_species_binomial)) +
    #if we desire NA to be the same as zero
    #https://blogs.uoregon.edu/rclub/2016/10/21/dealing-with-missingout-of-bounds-values-in-heatmaps/
    geom_tile(aes(x = x, y = y, fill = count))+
    scale_fill_viridis(option="mako", direction=-1,na.value = 'white')+
    theme_minimal()+
    theme(          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          legend.title = element_text(size = 6), 
               legend.text = element_text(size = 6),
          plot.margin = unit(c(0, 0,0,0),"inches"),
    )+
    #geom_sf(w)+
    labs(
      fill = "Number of species"
    )

ggplotly(p)%>% config(displayModeBar = T, scrollZoom =T, modeBarButtonsToRemove = buttonstoremove)%>% layout(dragmode="pan",
xaxis = list(range = list(0, 60)),
yaxis = list(range = list(0, 20))
      )
```

1a) Global species occurrence data were collected from the RBIEN database. Each species in the Kew World Checklist of Vascular Plants (WCVP) was searched in RBIEN (WCVP n=356,618, within RBIEN n= 174,106). Lat/long pairs (n=22,881,401) were downloaded and rasterized over a 4x4 latitude by longitude grid. 

## Figure 1b: Richness for vascular plant species with a sequenced transcriptome 



```{r fig1b, echo=FALSE, echo=FALSE, warning=FALSE, out.width='375px', out.height="667px"}
  rich_sra = raster("rich_just_sra_4.tif")
rwd <- as.data.frame(rich_sra, xy=T)
rwd$count <- rwd$scrubbed_species_binomial 

#fig3a <- plot_ly(rwd, x = ~x, y = ~y, z= ~count, type="heatmap", colors = viridis_pal(option = "G", direction=-1)(12))
#fig3a

p  <- ggplot(rwd, plot.margin = margin(0, 0, 0, 0, "cm")) +
   geom_sf(data=w, fill=NA)+
    #geom_spatraster(data = rich, aes(fill = scrubbed_species_binomial)) +
    #if we desire NA to be the same as zero
    #https://blogs.uoregon.edu/rclub/2016/10/21/dealing-with-missingout-of-bounds-values-in-heatmaps/
    geom_tile(aes(x = x, y = y, fill = count))+
    scale_fill_viridis(option="mako", direction=-1,na.value = 'white')+
    theme_minimal()+
    theme(          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          legend.title = element_text(size = 6), 
               legend.text = element_text(size = 6),
          plot.margin = unit(c(0, 0,0,0),"inches"),
    )+
    #geom_sf(w)+
    labs(
      fill = "Number of species"
    )

ggplotly(p)%>% config(displayModeBar = T, scrollZoom =T, modeBarButtonsToRemove = buttonstoremove, responsive = F)%>% layout(dragmode="pan",
xaxis = list(range = list(0, 60)),
yaxis = list(range = list(0, 20))
      )
```

1b) The records for transcriptome sequenced species which occurred in the SRA (n= 4,653 valid species were present in RBIEN, out of 6,966) were collected. This returned 6,441,980 lat/long pairs. 

## Figure 1c: Proportion of species with a sequenced transcriptome



```{r fig1c, echo=FALSE, echo=FALSE, warning=FALSE, out.width='375px', out.height="667px"}
    adj = raster("adj_rich_4.tif")
rwd <- as.data.frame(adj, xy=T)
rwd$count <- rwd$scrubbed_species_binomial 

#fig3a <- plot_ly(rwd, x = ~x, y = ~y, z= ~count, type="heatmap", colors = viridis_pal(option = "G", direction=-1)(12))
#fig3a

p  <- ggplot(rwd, plot.margin = margin(0, 0, 0, 0, "cm")) +
   geom_sf(data=w, fill=NA)+
    #geom_spatraster(data = rich, aes(fill = scrubbed_species_binomial)) +
    #if we desire NA to be the same as zero
    #https://blogs.uoregon.edu/rclub/2016/10/21/dealing-with-missingout-of-bounds-values-in-heatmaps/
    geom_tile(aes(x = x, y = y, fill = lyr1))+
    scale_fill_viridis(option="rocket", direction=-1,na.value = 'white')+
    theme_minimal()+
    theme(          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          legend.title = element_text(size = 6), 
               legend.text = element_text(size = 6),
          plot.margin = unit(c(0, 0,0,0),"inches"),
    )+
    #geom_sf(w)+
    labs(
      fill = "Proportion of species"
    )

ggplotly(p)%>% config(displayModeBar = T, scrollZoom =T, modeBarButtonsToRemove = buttonstoremove)%>% layout(dragmode="pan",
xaxis = list(range = list(0, 60)),
yaxis = list(range = list(0, 20))
      )
```

1c) An approximate measure of completeness was calculated by taking the number of species with transcriptomes sequenced, compared to the predicted number of species in a grid cell.